```{r}
#| label: calibration-prep
psm_clean <- janitor::clean_names(psm)

psm_ppm <- psm_clean %>%
  dplyr::mutate(
    Fraction = stringr::str_extract(spectrum_file, "[^\\\\/]+(?=\\.[^\\\\/]*$)"),
    ppm      = (delta_mass / calculated_peptide_mass) * 1e6
  ) %>%
  dplyr::select(Fraction, Peptide = peptide, Protein = protein, ppm)
```

## Calibration stability by fraction

Mass accuracy is expressed as the deviation between the measured and the theoretical peptide mass in parts per million (PPM):

$$\text{PPM} = \frac{\text{Delta Mass}}{\text{Calculated Peptide Mass}} \times 10^6$$

Values close to **0 PPM** indicate accurate mass calibration. A systematic shift affecting all fractions may reflect a global calibration offset, whereas fraction-specific deviations could point to instrument drift or sample-related issues during acquisition. 

<br>

### Summary
```{r}
#| label: ppm-stats-table
psm_ppm %>%
  dplyr::group_by(Fraction) %>%
  dplyr::summarise(
    N      = dplyr::n(),
    Median = round(median(ppm, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  dplyr::arrange(as.numeric(stringr::str_extract(Fraction, "\\d+$"))) %>%
  kableExtra::kbl(
    col.names = c("Fraction", "N", "Median (PPM)")
  ) %>%
  kableExtra::kable_paper("hover", full_width = FALSE, font_size = 15)
```

<br>

### Boxplot

```{r}
#| label: ppm-boxplot
#| fig-height: 5
#| fig-width: 9
ggplot(psm_ppm, aes(x = Fraction, y = ppm, fill = Fraction)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  coord_cartesian(ylim = c(-5, 5)) +
  theme_minimal() +
  labs(x = "Fraction",
       y = "Mass error (PPM)",
       title = "PPM deviation by fraction") +
  theme(
    axis.title.x = element_text(margin = margin(t = 15)),
    axis.title.y = element_text(margin = margin(r = 15)),
    axis.text.x  = element_text(angle = 90, vjust = 0.5),
    plot.title   = element_text(face = "bold", size = 11),
    legend.position = "none"
  )
```

### Histogram

```{r}
#| label: ppm-histogram
#| fig-height: 5
#| fig-width: 9
ggplot(psm_ppm, aes(x = ppm, fill = Fraction)) +
  geom_histogram(binwidth = 0.1, alpha = 0.7, position = "identity") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  coord_cartesian(xlim = c(-5, 5)) +
  theme_minimal() +
  labs(x = "Mass error (PPM)",
       y = "Count",
       title = "Mass error distribution") +
  theme(
    axis.title.x = element_text(margin = margin(t = 15)),
    axis.title.y = element_text(margin = margin(r = 15)),
    plot.title   = element_text(face = "bold", size = 11),
    legend.position = "none"
  )
```
