## Enrichment

### GO enrichment

The following figures accompany the excel file with more detailed information. For each comparison, the CC (cellular compartment), BP (biological process) and MF (molecular function) GO enriched terms are plotted (only those with an adjusted P value \< 0.05, or if none were found, by P value \< 0.05). If no figure is shown, no significant enrichment was found.

```{r}
df_imp = df_norm

list_prot_go <- list()

if (inherits(df_limma, "list")){
  
  for(i in 1:length(df_limma)){
    
    temp <- df_limma[[i]] %>% 
    left_join(dplyr::select(raw$protein,
                            Index,
                            Gene),
              by = c("Protein.IDs" = "Index"))
    
    if (nrow(temp[temp$P.Value < .05,]) == 0){
      
      cat("There were no diferentially expressed proteins in the comparison", 
          names(df_limma)[i], "\n")
      
    } else if (nrow(temp[which(temp$adj.P.Val < .05),]) > 0){
      
      ids_up <- temp$Protein.IDs[which(temp$adj.P.Val < .05 & temp$logFC > 0)]
      ids_down <- temp$Protein.IDs[which(temp$adj.P.Val < .05 & temp$logFC < 0)]
      
      ego_up <- enrich_prot2(ids_up, df_imp$Protein.IDs, org = org_key, readable = F)
      
      ego_down <- enrich_prot2(ids_down, df_imp$Protein.IDs, org = org_key, readable = F)
      
      list_prot_go[[i]] <- list("up" = ego_up, "down" = ego_down)
      names(list_prot_go)[i] <- names(df_limma)[i]
      
      
    } else if (nrow(temp[which(temp$P.Value < .05),]) > 0){
      
      ids_up <- temp$Protein.IDs[which(temp$P.Value < .05 & temp$logFC > 0)]
      ids_down <- temp$Protein.IDs[which(temp$P.Value < .05 & temp$logFC < 0)]
      
      ego_up <- enrich_prot2(ids_up, df_imp$Protein.IDs, org = org_key)
      
      ego_down <- enrich_prot2(ids_down, df_imp$Protein.IDs, org = org_key)
      
      list_prot_go[[i]] <- list("up" = ego_up, "down" = ego_down)
      names(list_prot_go)[i] <- names(df_limma)[i]
      
      }
    } 
  }

list_prot_go <- purrr::compact(purrr::list_flatten(list_prot_go))

names(list_prot_go) <- janitor::make_clean_names(names(list_prot_go))
```

```{r}
#| fig-align: center
#| fig-width: 10

figs <- list()

for (i in 1:length(list_prot_go)){
  
  if (nrow(list_prot_go[[i]]@result) > 0){
    
    temp <- list_prot_go[[i]]@result
    nam  <- names(list_prot_go)[i]
    
    direction <- dplyr::case_when(
      grepl("_up$",   nam) ~ "Up-regulated",
      grepl("_down$", nam) ~ "Down-regulated",
      TRUE                 ~ nam
    )
    
    color_title <- dplyr::case_when(
      grepl("_up$",   nam) ~ "#E64B35",
      grepl("_down$", nam) ~ "#4DBBD5",
      TRUE                 ~ "black"
    )
    
    orig_name <- gsub("_(up|down)$", "", nam)

    matched_key <- names(comp_labels)[
    janitor::make_clean_names(names(comp_labels)) == orig_name
    ]

    title_label <- paste0(comp_labels[matched_key], " — ", direction)
    
    figs[[i]] <- temp %>% 
      dplyr::mutate(Description = stringr::str_to_sentence(Description)) %>% 
      group_by(ONTOLOGY) %>% 
      slice(1:10) %>% 
      ungroup() %>% 
      ggplot(aes(FoldEnrichment, reorder(Description, FoldEnrichment))) +
      geom_point(aes(size = Count, color = p.adjust)) +
      labs(y     = "GO Term",
           size  = "Protein count",
           title = title_label) +
      theme_bw() +
      scale_color_gradient() +
      facet_grid(ONTOLOGY~., scale = "free") +
      theme(plot.title = element_text(color = color_title, face = "bold"))
    
    names(figs)[i] <- nam 
  }
}

figs_c <- purrr::compact(figs)

for (nm in names(figs_c)) {
  print(figs_c[[nm]])
}
```

```{r}
go_exp <- purrr::map(list_prot_go, ~.x@result) %>% 
  purrr::discard(~ nrow(.x) == 0)

openxlsx::write.xlsx(go_exp, file = paste0("results/",
                                              format(Sys.time(), "%Y%m%d"),
                                              exp_id,
                                              "_proteome_GO_enrichment.xlsx"))
```